---
title: "Chicago Crime Analysis"
output:
  html_document:
    df_print: paged
---

This is an analysis of crime rates in Chicago. Crime statistics were collected from [The Federal Bureau of Investigation](https://crime-data-explorer.fr.cloud.gov/pages/explorer/crime/crime-trend) and population data was collected from [FRED Economic Data](https://fred.stlouisfed.org/series/ILCOOK1POP).


# Data Prep

## Load packages
```{r}
library(tidyr)
library(ggplot2)
library(dplyr)
library(readr)
library(scales)
library(lubridate)
```





## [Population Data](https://fred.stlouisfed.org/series/ILCOOK1POP) 
### United States 
```{r}
# US Pop
Population_US <- read.csv("./Population/Population_US.csv")

# Create column called 'population'
Population_US$population <- Population_US$B230RC0A052NBEA

# Remove old columns
Population_US$B230RC0A052NBEA<-NULL

# Create Year Column for Join
Population_US$year<-substr(Population_US$DATE,0,4)

```

### Cook County (Chicago) 
```{r}
# Cook County Pop
Population_Cook <- read.csv("./Population/Population_CookCounty.csv")

# Create column called 'population'
Population_Cook$population <- Population_Cook$ILCOOK1POP

# Remove old columns
Population_Cook$ILCOOK1POP<-NULL

# Create Year Column for Join
Population_Cook$year<-substr(Population_Cook$DATE,0,4)

```

### Los Angeles 
```{r}
# Los Angeles County Pop
Population_LA <- read.csv("./Population/Population_LosAngeles.csv")

# Create column called 'population'
Population_LA$population <- Population_LA$CALOSA7POP

# Remove old columns
Population_LA$CALOSA7POP <-NULL

# Create Year Column for Join
Population_LA$year<-substr(Population_LA$DATE,0,4)

```

### New York City (5 boroughs) 
```{r}
# Pull in borough data
Population_Manhattan <- read.csv("./Population/Population_NYCounty.csv") #New York County = Manhattan
Population_Brooklyn <- read.csv("./Population/Population_Kings.csv") #Kings County = Brooklyn
Population_Bronx <- read.csv("./Population/Population_Bronx.csv") #Bronx County
Population_Staten <- read.csv("./Population/Population_Richmond.csv") #Richmond County = Staten Island
Population_Queens <- read.csv("./Population/Population_Queens.csv") #Queens County

Population_NYC<-left_join(Population_Manhattan,Population_Brooklyn,'DATE')
Population_NYC<-left_join(Population_NYC,Population_Bronx,'DATE')
Population_NYC<-left_join(Population_NYC,Population_Staten,'DATE')
Population_NYC<-left_join(Population_NYC,Population_Queens,'DATE')

# Create column called 'population'
Population_NYC<-Population_NYC %>% 
  group_by(DATE) %>% 
  summarize(population = NYNEWY1POP+NYKING7POP+NYBRON5POP+NYRICH5POP+NYQUEE1POP)

# Create Year Column for Join
Population_NYC$year<-substr(Population_NYC$DATE,0,4)

```





## [Crime Data](https://crime-data-explorer.fr.cloud.gov/pages/explorer/crime/crime-trend)
### United States Overall 
```{r}

# Create Tables for United States Stats (rates per 100K people per year)
AggravatedAssault <- read.csv("./Crime/AggravatedAssault.csv")
AllViolentCrime <- read.csv("./Crime/AllViolentCrime.csv")
Burglary <- read.csv("./Crime/Burglary.csv")
Homicide<- read.csv("./Crime/Homicide.csv")
MotorVehicleTheft <- read.csv("./Crime/MotorVehicleTheft.csv")
Robbery <- read.csv("./Crime/Robbery.csv")

# Tag Type in each dataset (facilitates transpose)
AggravatedAssault$Type <- 'Aggravated Assault'
AllViolentCrime$Type <- 'All Violent Crime'
Burglary$Type <- 'Burglary'
Homicide$Type <- 'Homicide'
MotorVehicleTheft$Type <- 'Motor Vehicle Theft'
Robbery$Type <- 'Robbery'

# Stack Data
US.df<-bind_rows(AggravatedAssault, AllViolentCrime,Burglary,Homicide,MotorVehicleTheft,Robbery)

# Pivot longer to create date column
US.df<-US.df %>% 
  pivot_longer(contains('x'),'year') %>% 
  mutate(year = gsub('X','',year),
         location = 'USA')

# Join on population Data
paste0('US has ',nrow(US.df),' rows prior to join')
US.df<-left_join(US.df, Population_US, 'year')
paste0('US has ',nrow(US.df),' rows after join')



```

### Chicago 
```{r}
# Chicago Specific (reported as raw number (not per 100K people))
AggravatedAssault_C <- read.csv("./Crime/AggravatedAssault_C.csv")
AllViolentCrime_C <- read.csv("./Crime/AllViolentCrime_C.csv")
Burglary_C <- read.csv("./Crime/Burglary_C.csv")
Homicide_C<- read.csv("./Crime/Homicide_C.csv")
MotorVehicleTheft_C <- read.csv("./Crime/MotorVehicleTheft_C.csv")
Robbery_C <- read.csv("./Crime/Robbery_C.csv")

# Tag Type in each dataset (facilitates transpose)
AggravatedAssault_C$Type <- 'Aggravated Assault'
AllViolentCrime_C$Type <- 'All Violent Crime'
Burglary_C$Type <- 'Burglary'
Homicide_C$Type <- 'Homicide'
MotorVehicleTheft_C$Type <- 'Motor Vehicle Theft'
Robbery_C$Type <- 'Robbery'

# Stack Data for Chicago (Cook County)
Cook.df<-bind_rows(AggravatedAssault_C, AllViolentCrime_C,Burglary_C,Homicide_C,MotorVehicleTheft_C,Robbery_C)

# Pivot longer to create date column
Cook.df<- Cook.df %>% 
  filter(series == 'Reported') %>% 
  pivot_longer(contains('x'),'year') %>% 
  mutate(year = gsub('X','',year),
         location = 'Chicago')

# Join on Population Data
paste0('Cook has ',nrow(Cook.df),' rows prior to join')
Cook.df<-left_join(Cook.df, Population_Cook, 'year')
paste0('Cook has ',nrow(Cook.df),' rows after join')

# Adjust Cook County's crimes to be listed as per 100K people (same as US Data)
Cook.df$value <- Cook.df$value/(Cook.df$population*1000/100000)
```

### LA 
```{r}
# LA Specific (reported as raw number (not per 100K people))
AggravatedAssault_LA <- read.csv("./Crime/AggravatedAssault_LA.csv")
AllViolentCrime_LA <- read.csv("./Crime/AllViolentCrime_LA.csv")
Burglary_LA <- read.csv("./Crime/Burglary_LA.csv")
Homicide_LA<- read.csv("./Crime/Homicide_LA.csv")
MotorVehicleTheft_LA <- read.csv("./Crime/MotorVehicleTheft_LA.csv")
Robbery_LA <- read.csv("./Crime/Robbery_LA.csv")

# Tag Type in each dataset (facilitates transpose)
AggravatedAssault_LA$Type <- 'Aggravated Assault'
AllViolentCrime_LA$Type <- 'All Violent Crime'
Burglary_LA$Type <- 'Burglary'
Homicide_LA$Type <- 'Homicide'
MotorVehicleTheft_LA$Type <- 'Motor Vehicle Theft'
Robbery_LA$Type <- 'Robbery'

# Stack Data for LA
LA.df<-bind_rows(AggravatedAssault_LA, AllViolentCrime_LA,Burglary_LA,Homicide_LA,MotorVehicleTheft_LA,Robbery_LA)

# Pivot longer to create date column
LA.df<- LA.df %>% 
  filter(series == 'Reported') %>% 
  pivot_longer(contains('x'),'year') %>% 
  mutate(year = gsub('X','',year),
         location = 'LA')

# Join on Population Data
paste0('LA has ',nrow(LA.df),' rows prior to join')
LA.df<-left_join(LA.df, Population_LA, 'year')
paste0('LA has ',nrow(LA.df),' rows after join')

# Adjust LA County's crimes to be listed as per 100K people (same as US Data)
LA.df$value <- LA.df$value/(LA.df$population*1000/100000)
```

### New York City (NYC) 
```{r}
# NYC Specific (reported as raw number (not per 100K people))
AggravatedAssault_NY <- read.csv("./Crime/AggravatedAssault_NY.csv")
AllViolentCrime_NY <- read.csv("./Crime/AllViolentCrime_NY.csv")
Burglary_NY <- read.csv("./Crime/Burglary_NY.csv")
Homicide_NY<- read.csv("./Crime/Homicide_NY.csv")
MotorVehicleTheft_NY <- read.csv("./Crime/MotorVehicleTheft_NY.csv")
Robbery_NY <- read.csv("./Crime/Robbery_NY.csv")

# Tag Type in each dataset (facilitates transpose)
AggravatedAssault_NY$Type <- 'Aggravated Assault'
AllViolentCrime_NY$Type <- 'All Violent Crime'
Burglary_NY$Type <- 'Burglary'
Homicide_NY$Type <- 'Homicide'
MotorVehicleTheft_NY$Type <- 'Motor Vehicle Theft'
Robbery_NY$Type <- 'Robbery'

# Stack Data for NYC
NYC.df<-bind_rows(AggravatedAssault_NY, AllViolentCrime_NY,Burglary_NY,Homicide_NY,MotorVehicleTheft_NY,Robbery_NY)

# Pivot longer to create date column
NYC.df<- NYC.df %>% 
  filter(series == 'Reported') %>% 
  pivot_longer(contains('x'),'year') %>% 
  mutate(year = gsub('X','',year),
         location = 'NYC')

# Join on Population Data
paste0('NYC has ',nrow(NYC.df),' rows prior to join')
NYC.df<-left_join(NYC.df, Population_NYC, 'year')
paste0('NYC has ',nrow(NYC.df),' rows after join')

# Adjust NYC crimes to be listed as per 100K people (same as US Data)
NYC.df$value <- NYC.df$value/(NYC.df$population*1000/100000)

```



## Stack Datasets & Remove 2021
```{r}
# Stack datasets
df <- bind_rows(US.df, Cook.df, NYC.df,LA.df)
df<-df %>%
  select(year, location, 'type' = Type, 'crimes.per' = value, population)  # Select headers
df$crimes<-df$crimes.per*(df$population*1000/100000) # (crimes per 100K people)x(population in hundred thousands)
                                                  # population is listed in thousands in the data
df<- df %>% 
  filter(year != '2021') #2021 data was incomplete for Chicago and not reported for LA & NYC
```






# Data Overview

## All Violent Crime: Chicago, LA, & NYC

### Graph: Total violent Crime Over Time
```{r, echo=FALSE}
# Total Violent Crime over time
df %>% 
  filter(type == "All Violent Crime") %>%
  filter(location != 'USA') %>% 
  group_by(location, year) %>% 
  summarize(crimes = sum(crimes)) %>% 
  ggplot(aes(x = as.integer(year), y = crimes, color = location))+
  geom_line()+
  theme_bw()+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  xlab('Year')+
  ylab('Violent Crimes')+
  ggtitle('Total Violent Crimes (1991-2020)')+
  scale_color_manual(values = c('NYC' = 'blue',
                                'Chicago' = 'red',
                                'LA' = 'lightblue3'))

```

### Table: Total Violent Crime Over Time
```{r}
# Table detailing All Violent Crime by year and city
table1<-df %>% 
  filter(location != "USA" & type == 'All Violent Crime' & year > '2016') %>% 
  group_by(location, year) %>% 
  summarize(crimes = sum(crimes)) %>%
  pivot_wider(names_from = year, values_from = crimes)

print(table1)
```


## Crimes per 100K People (crimes per capita)

### Graph: Violent Crime per Capita - Chicago, LA, & NYC
```{r, echo=FALSE}
# Total Violent Crime over time
df %>% 
  filter(type == "All Violent Crime") %>%
  group_by(location, year) %>% 
  summarize(crimes = sum(crimes.per)) %>% 
  ggplot(aes(x = as.integer(year), y = crimes, color = location))+
  geom_line()+
  theme_bw()+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  xlab('Year')+
  ylab('Violent Crimes per 100K People')+
  ggtitle('Total Violent Crimes per 100K People (1991-2020)')+
  scale_color_manual(values = c('NYC' = 'blue',
                                'Chicago' = 'red',
                                'LA' = 'lightblue3',
                                'USA' = 'black'))

```

### Facet Wrap: Past 3 years
```{r}
df %>% 
  filter(year > '2016') %>%   # Added filter for 2017-2020
  group_by(location, type, year) %>% 
  summarize(crimes = sum(crimes.per)) %>% 
  ggplot(aes(x = as.integer(year), y = crimes, color = location))+
  geom_line()+
  theme_bw()+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  facet_wrap(type~.)+
  ggtitle('Crimes per Capita by Location')+
  scale_color_manual(values = c('NYC' = 'blue2',
                                'Chicago' = 'red',
                                'LA' = 'lightblue3',
                                'USA' = 'black'))+
  ylab('Crimes per 100K People')+
  xlab('Year')
  
```



## [All Chicago Crime Data](https://data.cityofchicago.org/Public-Safety/Crimes-2001-to-Present/ijzp-q8t2)

### Clean File: All Chicago Crime
#### Load & filter Chicago source data 
```{r}

###### Full dataset from city of chicago (https://data.cityofchicago.org/Public-Safety/Crimes-2001-to-Present/ijzp-q8t2) ######
###### LONG RUN TIME. ONLY RUN IF NEED REFRESH. OTHERWISE USE 'ChicagoViolent.Post2017' ######

#ChicagoAll.df <- read.csv("./Chicago_All_Crime/Crimes_2001_to_Present.csv")
```

#### Write smaller file for use in analysis
```{r}

###### Filter full dataset to scope (violent crimes 2018 to present) ######

# ChicagoViolent.Post2017<-ChicagoAll.df %>%
#   select(Year, Case.Number, Date, Primary.Type, Description,
#          Location.Description, Domestic, District, Ward,
#          Community.Area, FBI.Code, Latitude, Longitude) %>%
#   filter(Year > 2017 &
#           !Primary.Type %in% c("CONCEALED CARRY LICENSE VIOLATION", "CRIMINAL DAMAGE", "CRIMINAL TRESPASS",
#                                 "DECEPTIVE PRACTICE", "GAMBLING", "INTERFERENCE WITH PUBLIC OFFICER",
#                                 "LIQUOR LAW VIOLATION", "NARCOTICS","NON-CRIMINAL", "STALKING",
#                                 "NON-CRIMINAL (SUBJECT SPECIFIED)", "OBSCENITY", "OFFENSE INVOLVING CHILDREN",
#                                 "OTHER NARCOTIC VIOLATION","OTHER OFFENSE","PROSTITUTION","PUBLIC INDECENCY",
#                                 "PUBLIC PEACE VIOLATION", "RITUALISM", "WEAPONS VIOLATION", "HUMAN TRAFFICKING"))


###### Create new dataset 'ChicagoViolent.Post2017'. Use this dataset for analysis going forward ######

# write.csv(ChicagoViolent.Post2017, "./Chicago_All_Crime/ChicagoPost2017.csv", row.names=FALSE)

```

#### Load new CSV File
```{r}
chi_violent.df <- read.csv("./Chicago_All_Crime/ChicagoPost2017.csv")
```


#### Name & Date Cleaning
```{r}
# Name Cleaning

chi_violent.df<- chi_violent.df %>% # Remove domestic crimes
  filter(Domestic != 'true')

chi_violent.df<-chi_violent.df %>% 
  mutate(Primary.Type.Clean = ifelse(Primary.Type == 'CRIM SEXUAL ASSAULT'
                        | Primary.Type == 'CRIMINAL SEXUAL ASSAULT'
                        | Primary.Type == 'SEX OFFENSE','SEXUAL', Primary.Type), # Sex crimes inconsistently tagged over time
         Primary.Type.Clean = ifelse(Primary.Type.Clean == 'KIDNAPPING'
                        | Primary.Type.Clean == 'INTIMIDATION'
                        | Primary.Type.Clean == 'ARSON'
                        | Primary.Type.Clean == 'SEXUAL','OTHER', Primary.Type.Clean)) # Group low occuring crimes
                                                                                      # Added sexual b/c low frequency

# Review Groupings
chi_violent.df %>% 
  group_by(Primary.Type.Clean, Primary.Type) %>% 
  summarise(Crimes = n()) %>% 
  arrange(Primary.Type.Clean)

# Add a count for each row (used to count individual case IDs)
chi_violent.df$count <- 1 

# Summarize date by first of month (month start)
chi_violent.df<-chi_violent.df %>% 
  mutate(month = substr(Date,0,2),
         month.start = paste0(Year,"-",month,"-01"))

# Convert month.start to date
chi_violent.df$month.start<-as.Date(chi_violent.df$month.start)

# Get day of month
chi_violent.df<-chi_violent.df %>%
  mutate( DateFull = Date, # Rename old Date Column
          day = substr(Date, 4,5), # Weekday
          Date = as.Date(paste0(Year,"-",month,"-",day)), #Create new Date column (easier to use)
          weekdayNum = wday(Date),
          time = substr(DateFull, 12, 23),
          hour = paste(substr(time,10,11),substr(time,0,2))) #Combine hour & AM/PM

# Assign names to weekdays
chi_violent.df$weekday <- recode(chi_violent.df$weekdayNum, 
                                 "1"="Sunday",
                                 "2"="Monday",
                                 "3"="Tuesday",
                                 "4"="Wednesday",
                                 "5"="Thursday",
                                 "6"="Friday",
                                 "7"="Saturday")

# Create 24hr format for graphs
chi_violent.df$hour24<-recode(chi_violent.df$hour, 
       "AM 12"= 00,
       "AM 01"= 01,
       "AM 02"= 02,
       "AM 03"= 03,
       "AM 04"= 04,
       "AM 05"= 05,
       "AM 06"= 06,
       "AM 07"= 07,
       "AM 08"= 08,
       "AM 09"= 09,
       "AM 10"= 10,
       "AM 11"= 11,
       "PM 12"= 12,
       "PM 01"= 13,
       "PM 02"= 14,
       "PM 03"= 15,
       "PM 04"= 16,
       "PM 05"= 17,
       "PM 06"= 18,
       "PM 07"= 19,
       "PM 08"= 20,
       "PM 09"= 21,
       "PM 10"= 22,
       "PM 11"= 23)
```
### Write Tableau File
```{r}
#write.csv(chi_violent.df, "./Tableau/ChicagoPost2017.csv", row.names=FALSE)
```







## Analyze all Chicago Violent Crimes since 2018
```{r}

# Summarize data by year and Primary.Type.Clean
annual.type<-chi_violent.df %>% 
  group_by(Year, Primary.Type.Clean) %>% 
  summarise(ByCategory = sum(count))

# Add annual totals
annual.type<-annual.type %>%
  group_by(Year) %>% 
  mutate(Total = sum(ByCategory),
         pct = round(ByCategory/Total,3)*100)

# Crimes - Total since 2018 by category
annual.type %>% 
  arrange(ByCategory) %>% 
  ggplot(aes(x = Year, y = ByCategory, color = Primary.Type.Clean))+
  geom_line()+
  theme_bw()+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  scale_color_brewer(palette = 'Accent')+
  ggtitle('Chicago Crimes: Total by Category since 2018')

# Crimes - Percent of Total
annual.type %>% 
  arrange(ByCategory) %>% 
  ggplot(aes(x = Year, y = pct, fill = Primary.Type.Clean, label = pct))+
  geom_bar(stat = 'identity')+
  theme_bw()+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  scale_fill_brewer(palette = 'Accent')+
  geom_text(size = 3, position = position_stack(vjust = 0.5))+
  ggtitle('Chicago Crimes: Percent of total since 2018')

# Crimes - Total  
for (i in unique(annual.type$Primary.Type.Clean)){

chart<-annual.type %>%
  filter(Primary.Type.Clean == i) %>% 
  arrange(ByCategory) %>% 
  ggplot(aes(x = Year, y = ByCategory))+
  geom_col()+
  theme_bw()+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  scale_fill_brewer(palette = 'Accent')+
  ggtitle(i,'since 2018')

print(chart)
}

```

### Heat Maps: Months & Years

```{r}
for (i in unique(chi_violent.df$Primary.Type.Clean)){

heat<-chi_violent.df %>%
  filter(Primary.Type.Clean == i) %>% 
  group_by(Year, month) %>% 
  summarise(crimes = sum(count)) %>% 
  ggplot(aes(x = Year, y = month, fill = -crimes))+
  geom_tile()+
  geom_text(aes(label = format(round(crimes,0))))+
  ggtitle(paste(i))

print(heat)

}
```
### Heatmaps: Days & Times
```{r}

for (i in unique(chi_violent.df$Primary.Type.Clean)){

heat2<-chi_violent.df %>%
  filter(Primary.Type.Clean == i) %>% 
  group_by(weekdayNum, hour24) %>% 
  summarise(crimes = sum(count)) %>% 
  ggplot(aes(x = weekdayNum, y = hour24, fill = desc(crimes)))+
  geom_tile()+
  geom_text(aes(label = format(round(crimes,0))))+
  ggtitle(i)+
  scale_x_continuous(breaks = seq(1,7, by = 1))+
  scale_y_continuous(breaks = seq(0,24, by = 2))

print(heat2)
 
}
```

### Word Clouds

#### Word Cloud: 2022 Theft
```{r}
library(wordcloud2)
library(RColorBrewer)
library(wordcloud2)
library(tm)


# Word cloud for locations of theft in 2022
descriptions <- chi_violent.df %>% 
  filter(Primary.Type.Clean == 'THEFT' & Year == '2022') %>% 
  select(Location.Description)

descriptions<-gsub(' LOT','LOT', descriptions)
docs <- Corpus(VectorSource(descriptions))

docs <- docs %>%
  tm_map(removeNumbers) %>%
  tm_map(removePunctuation) %>%
  tm_map(stripWhitespace)
docs <- tm_map(docs, content_transformer(tolower))
docs <- tm_map(docs, removeWords, stopwords("english"))

dtm <- TermDocumentMatrix(docs) 
matrix <- as.matrix(dtm) 
words <- sort(rowSums(matrix),decreasing=TRUE) 
df.T <- data.frame(word = names(words),freq=words)

set.seed(1234) # for reproducibility 
wordcloud2(df.T)
```

#### Word Cloud: 2022 Motor Vehicle Theft
```{r}

# Word cloud for locations of theft in 2022
descriptions <- chi_violent.df %>% 
  filter(Primary.Type.Clean == 'MOTOR VEHICLE THEFT' & Year == '2022') %>% 
  select(Location.Description)

descriptions<-gsub(' LOT','LOT', descriptions)
docs <- Corpus(VectorSource(descriptions))

docs <- docs %>%
  tm_map(removeNumbers) %>%
  tm_map(removePunctuation) %>%
  tm_map(stripWhitespace)
docs <- tm_map(docs, content_transformer(tolower))
docs <- tm_map(docs, removeWords, stopwords("english"))

dtm <- TermDocumentMatrix(docs) 
matrix <- as.matrix(dtm) 
words <- sort(rowSums(matrix),decreasing=TRUE) 
df.MVT <- data.frame(word = names(words),freq=words)

set.seed(1234) # for reproducibility 
wordcloud2(df.MVT)
```
#### Word Cloud: 2022 BATTERY
```{r}

# Word cloud for locations of theft in 2022
descriptions <- chi_violent.df %>% 
  filter(Primary.Type.Clean == 'BATTERY' & Year == '2022') %>% 
  select(Location.Description)

descriptions<-gsub(' LOT','LOT', descriptions)
docs <- Corpus(VectorSource(descriptions))

docs <- docs %>%
  tm_map(removeNumbers) %>%
  tm_map(removePunctuation) %>%
  tm_map(stripWhitespace)
docs <- tm_map(docs, content_transformer(tolower))
docs <- tm_map(docs, removeWords, stopwords("english"))

dtm <- TermDocumentMatrix(docs) 
matrix <- as.matrix(dtm) 
words <- sort(rowSums(matrix),decreasing=TRUE) 
df.B <- data.frame(word = names(words),freq=words)

set.seed(1234) # for reproducibility 
wordcloud2(df.B)
```

